Internal models of type theory and internal higher categories for the Univalent Foundations

### 2014-02-26 article-models-tt.tex
lots of work on extension on terms

\begin{align}
&\inference{\jfam{\ctxext{{\Gamma}{A}}{P}}{Q}}
{\jfameq{\Gamma}{\subst{x}{{f}{Q}}}{\subst{{x}{f}}{{x}{Q}}}}\\
&\inference{\jfam{\ctxext{{{\Gamma}{A}}{P}}{Q}}{R}}
{\jfameq{\ctxext{\Gamma}{\subst{x}{{f}{Q}}}}{\subst{x}{{f}{R}}}{\subst{{x}{f}}{{x}{R}}}}\\
&\inference{\jterm{\ctxext{{\Gamma}{A}}{P}}{Q}{g}}
{\jtermeq{\Gamma}{\subst{x}{{f}{Q}}}{\subst{x}{{f}{g}}}{\subst{{x}{f}}{{x}{g}}}}\\
&\inference{\jterm{\ctxext{{{\Gamma}{A}}{P}}{Q}}{R}{h}}
{\jtermeq{\ctxext{\Gamma}{\subst{x}{{f}{Q}}}}{\subst{x}{{f}{R}}}{\subst{x}{{f}{h}}}{\subst{{x}{f}}{{x}{h}}}}
\end{align}

@@

\subsection{Extension on terms}\label{extension-on-terms}
When $\jterm{\Gamma}{A}{x}$ and $\jterm{\Gamma}{\subst{x}{P}}{u}$ are terms,
we obtain a term $\jterm{\Gamma}{\ctxext{A}{P}}{\ctxext{x}{u}}$. This operation
of weakening on terms is like the pairing operation and like with the pairing
operation, extension on terms tells what the terms of an extended family are:
every term of an extended family decomposes into the components of a pair. 
\begin{align}
& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}}{\jterm{\Gamma}{\ctxext{A}{P}}{\ctxext{x}{u}}}
& & \inference{\jtermeq{\Gamma}{A}{x}{x'}\quad\jtermeq{\Gamma}{\subst{x}{P}}{u}{u'}}{\jtermeq{\Gamma}{\ctxext{A}{P}}{\ctxext{x}{u}}{\ctxext{x'}{u'}}}\\
& \inference{\jterm{\Gamma}{\ctxext{A}{P}}{w}}{\jterm{\Gamma}{A}{\pts{w}}}
& & \inference{\jtermeq{\Gamma}{\ctxext{A}{P}}{w}{w'}}{\jtermeq{\Gamma}{A}{\pts{w}}{\pts{w'}}}\\
& \inference{\jterm{\Gamma}{\ctxext{A}{P}}{w}}{\jterm{\Gamma}{\subst{\pts{w}}{P}}{\edg{w}}}
& & \inference{\jtermeq{\Gamma}{\ctxext{A}{P}}{w}{w'}}{\jtermeq{\Gamma}{\subst{\pts{w}}{P}}{\edg{w}}{\edg{w'}}}
\end{align}
However, this pairing is handled
much stricter than $\smsym$-types, 
because the decomposition is a judgmental inverse (and it will be subject to various compatibility
rules).
\begin{align}
& \inference{\jterm{\Gamma}{\ctxext{A}{P}}{w}}{\jtermeq{\Gamma}{\ctxext{A}{P}}{\ctxext{\pts{w}}{\edg{w}}}{w}}\\
& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}}{\jtermeq{\Gamma}{A}{\pts{(\ctxext{x}{u})}}{x}}\\
& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}}{\jtermeq{\Gamma}{\subst{x}{P}}{\edg{(\ctxext{x}{u})}}{u}}
\end{align}

We add the following compatibility rule for two consecutive term extensions:
\begin{equation}
\inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}\quad\jterm{\Gamma}{\subst{\ctxext{x}{u}}{Q}}{v}}
{\jtermeq{\Gamma}{\ctxext{{A}{P}}{Q}}{\ctxext{x}{{u}{v}}}{\ctxext{{x}{u}}{v}}}
\end{equation}

When we weaken a term $\ctxext{y}{v}$ of $\ctxext{B}{Q}$ in context $\Gamma$ by
a type $A$, the term that we get is $\ctxext{\ctxwk{A}{y}}{\ctxwk{A}{v}}$.
\begin{align}
& \inference{\jterm{\Gamma}{B}{y}\quad\jterm{\Gamma}{\subst{y}{Q}}{v}}{\jtermeq{\ctxext{\Gamma}{A}}{\ctxwk{A}{\ctxext{B}{Q}}}{\ctxwk{A}{\ctxext{y}{v}}}{\ctxext{\ctxwk{A}{y}}{\ctxwk{A}{v}}}}\\
& \inference{\jterm{\ctxext{\Gamma}{B}}{Q}{g}\quad\jterm{\ctxext{\Gamma}{B}}{\subst{g}{R}}{t}}{\jtermeq{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{\ctxext{Q}{R}}}{\ctxwk{A}{\ctxext{g}{t}}}{\ctxext{\ctxwk{A}{g}}{\ctxwk{A}{t}}}}
\end{align}

When we substitute an extended term $\ctxext{f}{g}$ of $\ctxext{P}{Q}$ by a term
$x$ of $A$, the term that we get is $\ctxext{\subst{x}{f}}{\subst{x}{g}}$.
\begin{align}
& \inference{\jterm{\ctxext{\Gamma}{A}}{P}{f}\quad\jterm{\ctxext{\Gamma}{A}}{\subst{f}{Q}}{g}}{\jtermeq{\Gamma}{\ctxext{\subst{x}{P}}{\subst{x}{Q}}}{\subst{x}{\ctxext{f}{g}}}{\ctxext{\subst{x}{f}}{\subst{x}{g}}}}\\
& \inference{\jterm{\ctxext{{\Gamma}{A}}{P}}{Q}{g}\quad\jterm{\ctxext{{\Gamma}{A}}{P}}{\subst{g}{R}}{t}}{\jtermeq{\ctxext{\Gamma}{\subst{x}{P}}}{\ctxext{\subst{x}{Q}}{\subst{x}{R}}}{\subst{x}{\ctxext{g}{t}}}{\ctxext{\subst{x}{g}}{\subst{x}{t}}}}
\end{align}

Finally, when we substitute by an extended term we get an equal result as when we
substitute two consecutive times. This looks like the Curry-Howard isomorphism.
\begin{align}
& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}\quad\jfam{\ctxext{{\Gamma}{A}}{P}}{Q}}
  {\jfameq{\Gamma}{\subst{\ctxext{x}{u}}{Q}}{\subst{u}{{x}{Q}}}}\\
& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}\quad\jterm{\ctxext{{\Gamma}{A}}{P}}{Q}{g}}
  {\jtermeq{\Gamma}{\subst{u}{{x}{Q}}}{\subst{\ctxext{x}{u}}{g}}{\subst{u}{{x}{g}}}}\\
& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}\quad\jfam{\ctxext{{{\Gamma}{A}}{P}}{Q}}{R}}
  {\jfameq{\ctxext{\Gamma}{\subst{u}{{x}{Q}}}}{\subst{\ctxext{x}{u}}{R}}{\subst{u}{{x}{R}}}}\\
& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}\quad\jterm{\ctxext{{{\Gamma}{A}}{P}}{Q}}{R}{t}}
  {\jtermeq{\ctxext{\Gamma}{\subst{u}{{x}{Q}}}}{\subst{u}{{x}{R}}}{\subst{\ctxext{x}{u}}{t}}{\subst{u}{{x}{t}}}}
\end{align}

@@

\begin{align}
& \inference{\jfam{\Gamma}{A}\qquad\jfam{\Gamma}{B}\qquad\jterm{\Gamma}{A}{x}}{\jfameq{\Gamma}{\subst{x}{\ctxwk{A}{B}}}{B}}\\
& \inference{\jterm{\Gamma}{A}{x}\qquad\jterm{\Gamma}{B}{y}}{\jtermeq{\Gamma}{B}{\subst{x}{\ctxwk{A}{y}}}{y}}\\
& \inference{\jfam{\Gamma}{A}\qquad\jfam{\ctxext{\Gamma}{B}}{Q}\qquad\jterm{\Gamma}{A}{x}}{\jfameq{\ctxext{\Gamma}{B}}{\subst{x}{\ctxwk{A}{Q}}}{Q}}\\
& \inference{\jterm{\Gamma}{A}{x}\qquad\jterm{\ctxext{\Gamma}{B}}{Q}{g}}{\jtermeq{\ctxext{\Gamma}{B}}{Q}{\subst{x}{\ctxwk{A}{g}}}{g}}
\end{align}

@@

\subsubsection{Identity functions}
Without a rule explicitly asserting the existence of an identity morphism we don't
get one, hence we do that here. The identity morphism is a term which introduced
in ordinary type theory via the variable rule. The variable rule is a bit more
general: it asserts that
\begin{equation*}
\jterm{\Gamma,\,x_1:A_1,\ldots,\,x_n:A_n}{A_i}{x_i}
\end{equation*}
for every $1\leq i\leq n$. Thus, it establishes the projections. In our setting,
we get the projections from the identity morphisms together with weakening. We
already have weakening, so here it suffices to introduce the identity morphisms.
\begin{align}
& \inference{\jfam{\Gamma}{A}}{\jterm{\ctxext{\Gamma}{A}}{\ctxwk{A}{A}}{\idfunc[A]}}
& & \inference{\jfameq{\Gamma}{A}{A'}}{\jtermeq{\ctxext{\Gamma}{A}}{\ctxwk{A}{A}}{\idfunc[A]}{\idfunc[A']}}
\end{align}
Identity functions are determined by their behavior with respect to substitution combined with
weakening. The identity functions will also be subject to compatibility rules.
\begin{align}
& \inference{\jterm{\Gamma}{A}{x}}{\jtermeq{\Gamma}{A}{\subst{x}{\idfunc[A]}}{x}}\\
& \inference{\jfam{\ctxext{\Gamma}{A}}{P}}{\jfameq{\ctxext{\Gamma}{A}}{\subst{\idfunc[A]}{\ctxwk{A}{P}}}{P}}\\
& \inference{\jterm{\ctxext{\Gamma}{A}}{P}{f}}{\jtermeq{\ctxext{\Gamma}{A}}{P}{\subst{\idfunc[A]}{\ctxwk{A}{f}}}{f}} \\
& \inference{\jterm{\ctxext{\Gamma}{B}}{\ctxwk{B}{A}}{g}}{\jtermeq{\ctxext{\Gamma}{B}}{\ctxwk{B}{A}}{\subst{g}{\ctxwk{B}{\idfunc[A]}}}{g}}
\end{align}

Using these rules we can derive that there is a judgmental equality 
$\jfameq{\Gamma}{\subst{x}{{\idfunc[A]}{R}}}
{\subst{x}{{x}{R}}}$ for any $\jfam{\ctxext{{\Gamma}{A}}{\ctxwk{A}{A}}}{R}$. 
Note that $\subst{x}{{x}{R}}$ is judgmentatlly the same as
$\subst{\ctxext{x}{x}}{R}$.

The identity functions in combination with weakening give rise to the projections.
We define the term
\begin{equation*}
\proj P \defeq \ctxwk{P}{\idfunc[A]}:\ctxwk{P}{{A}{A}}
\end{equation*}
in context $\ctxext{{\Gamma}{A}}{P}$.

The identity function of an extended family is the extension of the identity
functions on the components in the apropriate way. To find out what the
apropriate way is, note that
\begin{align*}
\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}} & \jdeq \ctxwk{P}{{A}{\ctxext{A}{P}}}\\
& \jdeq \ctxext{\ctxwk{P}{{A}{A}}}{\ctxwk{P}{{A}{P}}}
\end{align*}
We have the term $\jterm{\ctxext{\Gamma}{{A}{P}}}{\ctxwk{P}{A}}{\ctxwk{P}{\idfunc[A]}}$.
Thus we need to find out what $\subst{\ctxwk{P}{\idfunc[A]}}{\ctxwk{P}{{A}{P}}}$ is:
\begin{align*}
\subst{\ctxwk{P}{\idfunc[A]}}{\ctxwk{P}{{A}{P}}} & \jdeq \ctxwk{P}{\subst{\idfunc[A]}{\ctxwk{A}{P}}}\\
& \jdeq \ctxwk{P}{P},
\end{align*}
where we find the term $\idfunc[P]$. Therefore, we have the following inference rule:
\begin{equation}
\inference{\jfam{\Gamma}{A}\quad\jfam{\ctxext{\Gamma}{A}}{P}}{\jtermeq{\ctxext{\Gamma}{{A}{P}}}{\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}}{\idfunc[\ctxext{A}{P}]}{\ctxext{\ctxwk{P}{\idfunc[A]}}{\idfunc[P]}}}
\end{equation}

The identity function of a weakened family is the weakened identity function:
\begin{equation}
\inference{\jfam{\Gamma}{A}\quad\jfam{\Gamma}{B}}{\jtermeq{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{B}}}{\idfunc[\ctxwk{A}{B}]}{\ctxwk{A}{\idfunc[B]}}}
\end{equation}

The identity function of a substituted family is the substitution of the identity function
\begin{equation}
\inference{\jterm{\Gamma}{A}{x}\quad\jfam{\ctxext{\Gamma}{A}}{P}}{\jtermeq{\ctxext{\Gamma}{\subst{x}{P}}}{\ctxwk{\subst{x}{P}}{\subst{x}{P}}}{\idfunc[\subst{x}{P}]}{\subst{x}{\idfunc[P]}}}
\end{equation}

>>>>

\begin{align}
&\inference{\jfam{\ctxext{{\Gamma}{A}}{P}}{Q}}
{\jfameq{\Gamma}{\subst{x}{{f}{Q}}}{\subst{{x}{f}}{{x}{Q}}}}\label{comp-ss-1}\\
&\inference{\jterm{\ctxext{{\Gamma}{A}}{P}}{Q}{g}}
{\jtermeq{\Gamma}{\subst{x}{{f}{Q}}}{\subst{x}{{f}{g}}}{\subst{{x}{f}}{{x}{g}}}}\label{comp-ss-2}\\
&\inference{\jfam{\ctxext{{{\Gamma}{A}}{P}}{Q}}{R}}
{\jfameq{\ctxext{\Gamma}{\subst{x}{{f}{Q}}}}{\subst{x}{{f}{R}}}{\subst{{x}{f}}{{x}{R}}}}\label{comp-ss-3}\\
&\inference{\jterm{\ctxext{{{\Gamma}{A}}{P}}{Q}}{R}{h}}
{\jtermeq{\ctxext{\Gamma}{\subst{x}{{f}{Q}}}}{\subst{x}{{f}{R}}}{\subst{x}{{f}{h}}}{\subst{{x}{f}}{{x}{h}}}}\label{comp-ss-4}
\end{align}

@@

@@

\begin{align}
& \inference{\jfam{\Gamma}{A}\qquad\jfam{\Gamma}{B}\qquad\jterm{\Gamma}{A}{x}}{\jfameq{\Gamma}{\subst{x}{\ctxwk{A}{B}}}{B}}\label{defn-ws-1}\\
& \inference{\jterm{\Gamma}{A}{x}\qquad\jterm{\Gamma}{B}{y}}{\jtermeq{\Gamma}{B}{\subst{x}{\ctxwk{A}{y}}}{y}}\label{defn-ws-2}\\
& \inference{\jfam{\Gamma}{A}\qquad\jfam{\ctxext{\Gamma}{B}}{Q}\qquad\jterm{\Gamma}{A}{x}}{\jfameq{\ctxext{\Gamma}{B}}{\subst{x}{\ctxwk{A}{Q}}}{Q}}\label{defn-ws-3}\\
& \inference{\jterm{\Gamma}{A}{x}\qquad\jterm{\ctxext{\Gamma}{B}}{Q}{g}}{\jtermeq{\ctxext{\Gamma}{B}}{Q}{\subst{x}{\ctxwk{A}{g}}}{g}}\label{defn-ws-4}
\end{align}

@@

\subsubsection{Identity functions}
Without a rule explicitly asserting the existence of an identity morphism we don't
get one, hence we do that here. The identity morphism is a term which introduced
in ordinary type theory via the variable rule. The variable rule is a bit more
general: it asserts that
\begin{equation*}
\jterm{\Gamma,\,x_1:A_1,\ldots,\,x_n:A_n}{A_i}{x_i}
\end{equation*}
for every $1\leq i\leq n$. Thus, it establishes the projections. In our setting,
we get the projections from the identity morphisms together with weakening. We
already have weakening, so here it suffices to introduce the identity morphisms.
\begin{align}
& \inference{\jfam{\Gamma}{A}}{\jterm{\ctxext{\Gamma}{A}}{\ctxwk{A}{A}}{\idfunc[A]}}
& & \inference{\jfameq{\Gamma}{A}{A'}}{\jtermeq{\ctxext{\Gamma}{A}}{\ctxwk{A}{A}}{\idfunc[A]}{\idfunc[A']}}
\end{align}
Identity functions are determined by their behavior with respect to substitution combined with
weakening. The identity functions will also be subject to compatibility rules.
\begin{align}
& \inference{\jterm{\Gamma}{A}{x}}{\jtermeq{\Gamma}{A}{\subst{x}{\idfunc[A]}}{x}}\label{idfunc-subst-defn}\\
& \inference{\jfam{\ctxext{\Gamma}{A}}{P}}{\jfameq{\ctxext{\Gamma}{A}}{\subst{\idfunc[A]}{\ctxwk{A}{P}}}{P}}\label{idfunc-wk-defn}\\
& \inference{\jterm{\ctxext{\Gamma}{A}}{P}{f}}{\jtermeq{\ctxext{\Gamma}{A}}{P}{\subst{\idfunc[A]}{\ctxwk{A}{f}}}{f}}\label{idfunc-precomp} \\
& \inference{\jterm{\ctxext{\Gamma}{B}}{\ctxwk{B}{A}}{g}}{\jtermeq{\ctxext{\Gamma}{B}}{\ctxwk{B}{A}}{\subst{g}{\ctxwk{B}{\idfunc[A]}}}{g}}\label{idfunc-postcomp}
\end{align}

The identity function of an extended family is the extension of the identity
functions on the components in the apropriate way. To find out what the
apropriate way is, note that
\begin{align*}
\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}} & \jdeq \ctxwk{P}{{A}{\ctxext{A}{P}}}\\
& \jdeq \ctxext{\ctxwk{P}{{A}{A}}}{\ctxwk{P}{{A}{P}}}
\end{align*}
We have the term $\jterm{\ctxext{\Gamma}{{A}{P}}}{\ctxwk{P}{A}}{\ctxwk{P}{\idfunc[A]}}$.
Thus we need to find out what $\subst{\ctxwk{P}{\idfunc[A]}}{\ctxwk{P}{{A}{P}}}$ is:
\begin{align*}
\subst{\ctxwk{P}{\idfunc[A]}}{\ctxwk{P}{{A}{P}}} & \jdeq \ctxwk{P}{\subst{\idfunc[A]}{\ctxwk{A}{P}}}\\
& \jdeq \ctxwk{P}{P},
\end{align*}
where we find the term $\idfunc[P]$. Therefore, we have the following inference rule:
\begin{equation}\label{idfunc-ext-comp}
\inference{\jfam{\Gamma}{A}\quad\jfam{\ctxext{\Gamma}{A}}{P}}{\jtermeq{\ctxext{\Gamma}{{A}{P}}}{\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}}{\idfunc[\ctxext{A}{P}]}{\subst{\idfunc[P]}{{\ctxwk{P}{\idfunc[A]}}{\idfunc[\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}]}}}}
\end{equation}

The identity function of a weakened family is the weakened identity function:
\begin{equation}\label{idfunc-wk-comp}
\inference{\jfam{\Gamma}{A}\quad\jfam{\Gamma}{B}}{\jtermeq{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{B}}}{\idfunc[\ctxwk{A}{B}]}{\ctxwk{A}{\idfunc[B]}}}
\end{equation}

The identity function of a substituted family is the substitution of the identity function
\begin{equation}\label{idfunc-subst-comp}
\inference{\jterm{\Gamma}{A}{x}\quad\jfam{\ctxext{\Gamma}{A}}{P}}{\jtermeq{\ctxext{\Gamma}{\subst{x}{P}}}{\ctxwk{\subst{x}{P}}{\subst{x}{P}}}{\idfunc[\subst{x}{P}]}{\subst{x}{\idfunc[P]}}}
\end{equation}

\subsection{Extension on terms}\label{extension-on-terms}
When $\jterm{\Gamma}{A}{x}$ and $\jterm{\Gamma}{\subst{x}{P}}{u}$ are terms,
we obtain the term $\jterm{\Gamma}{\ctxext{A}{P}}{\subst{u}{{x}{\idfunc[\ctxext{A}{P}]}}}$. 
We define 
\begin{equation*}
\ctxext{x}{u}\defeq\subst{u}{{x}{\idfunc[\ctxext{A}{P}]}}.
\end{equation*}
Thus
$\ctxext{x}{u}$ is the pairing of $x$ and $u$. Note that because
$\ctxwk{P}{{A}{\ctxext{A}{P}}}\jdeq\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}$, the
pairing function is just $\idfunc[\ctxext{A}{P}]$.

There is also a decomposition of terms $\jterm{\Gamma}{\ctxext{A}{P}}{w}$ into
components: the projections. We define
\begin{align*}
\cprojfstf{A}{P} & \defeq \ctxwk{P}{\idfunc[A]}\\
\cprojsndf{A}{P} & \defeq \idfunc[P]
\end{align*}

The constructions of the terms $\ctxext{x}{u}$ and $\cprojfst{A}{P}{w}$ and
$\cprojsnd{A}{P}{w}$ are subject to various rules, with all of them being
consequences of earlier introduced inference rules.

\begin{lem} The following inference rules expressing that pairing is a strict
inverse to the combination of decompositions, are valid:
\begin{align*}
& \inference{\jterm{\Gamma}{\ctxext{A}{P}}{w}}{\jtermeq{\Gamma}{\ctxext{A}{P}}{\ctxext{\cprojfst{A}{P}{w}}{\cprojsnd{A}{P}{w}}}{w}}\\
& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}}{\jtermeq{\Gamma}{A}{\cprojfst{A}{P}{\ctxext{x}{u}}}{x}}\\
& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}}{\jtermeq{\Gamma}{\subst{x}{P}}{\cprojsnd{A}{P}{\ctxext{x}{u}}}{u}}
\end{align*}
\end{lem}

\begin{proof}
To prove the first judgmental equality, note that
\begin{align*}
w & \jdeq \subst{w}{\idfunc[\ctxext{A}{P}]} \tag{by \autoref{idfunc-subst-defn}}\\
& \jdeq \subst{w}{\subst{\idfunc[P]}{{\ctxwk{P}{\idfunc[A]}}{\idfunc[\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}]}}}
  \tag{by \autoref{idfunc-ext-comp}}\\
& \jdeq \subst{{w}{\idfunc[P]}}{\subst{w}{{\ctxwk{P}{\idfunc[A]}}{\idfunc[\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}]}}}\tag{by \autoref{comp-ss-4}}\\
& \jdeq \subst{{w}{\idfunc[P]}}{{\subst{w}{\ctxwk{P}{\idfunc[A]}}}{\subst{w}{\idfunc[\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}]}}}\tag{by \autoref{comp-ss-2}}\\
& \jdeq \subst{{w}{\idfunc[P]}}{{\subst{w}{\ctxwk{P}{\idfunc[A]}}}{\subst{w}{\ctxwk{\ctxext{A}{P}}{\idfunc[\ctxext{A}{P}]}}}}\tag{by \autoref{idfunc-wk-comp}}\\
& \jdeq \subst{{w}{\idfunc[P]}}{{\subst{w}{\ctxwk{P}{\idfunc[A]}}}{\idfunc[\ctxext{A}{P}]}}\tag{by \autoref{defn-ws-2}}\\
& \jdeq \ctxext{\cprojfst{A}{P}{w}}{\cprojsnd{A}{P}{w}}\tag{by definition}
\end{align*}
\end{proof}

\begin{lem}
The following compatibility rules for two consecutive term extensions is valid:
\begin{align*}
& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}\quad\jterm{\Gamma}{\subst{\ctxext{x}{u}}{Q}}{v}}
{\jtermeq{\Gamma}{\ctxext{{A}{P}}{Q}}{\ctxext{x}{{u}{v}}}{\ctxext{{x}{u}}{v}}}\\
& \inference{\jterm{\Gamma}{\ctxext{{A}{P}}{Q}}{w}}
  {\jtermeq{\Gamma}{A}{\cprojfst{A}{P}{\cprojfst{\ctxext{A}{P}}{Q}{w}}}{\cprojfst{A}{\ctxext{P}{Q}}{w}}}\\
& \inference{\jterm{\Gamma}{\ctxext{{A}{P}}{Q}}{w}}
  {\jtermeq{\Gamma}{\subst{\cprojfst{A}{\ctxext{P}{Q}}{w}}{P}}
  {\cprojsnd{A}{P}{\cprojfst{\ctxext{A}{P}}{Q}{w}}}
  {\cprojfst{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}}\\
& \inference{\jterm{\Gamma}{\ctxext{{A}{P}}{Q}}{w}}
  {\jtermeq{\Gamma}{\subst{\cprojfst{\ctxext{A}{P}}{Q}{w}}{Q}}
  {\cprojsnd{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}
  {\cprojsnd{\ctxext{A}{P}}{Q}{w}}}
\end{align*}
\end{lem}

\begin{lem}
When we weaken a term $\ctxext{y}{v}$ of $\ctxext{B}{Q}$ in context $\Gamma$ by
a type $A$, the term that we get is $\ctxext{\ctxwk{A}{y}}{\ctxwk{A}{v}}$. More
precisely, the following inference rules are valid:
\begin{align*}
& \inference{\jterm{\ctxext{\Gamma}{B}}{Q}{g}\quad\jterm{\ctxext{\Gamma}{B}}{\subst{g}{R}}{t}}{\jtermeq{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{\ctxext{Q}{R}}}{\ctxwk{A}{\ctxext{g}{t}}}{\ctxext{\ctxwk{A}{g}}{\ctxwk{A}{t}}}}
% \\
%& \inference{\jterm{\Gamma}{\ctxwk{A}{\ctxext{B}{Q}}}{\ctxwk{A}{w}}}{\jtermeq{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}{\cprojfst{\ctxwk{A}{B}}{\ctxwk{A}{Q}}{\ctxwk{A}{w}}}{\ctxwk{A}{(\cprojfst{B}{Q}{w})}}}\\
\end{align*}
\end{lem}

\begin{lem}
When we substitute an extended term $\ctxext{f}{g}$ of $\ctxext{P}{Q}$ by a term
$x$ of $A$, the term that we get is $\ctxext{\subst{x}{f}}{\subst{x}{g}}$.
More precisely, the following inference rules are valid:
\begin{align*}
& \inference{\jterm{\ctxext{\Gamma}{A}}{P}{f}\quad\jterm{\ctxext{\Gamma}{A}}{\subst{f}{Q}}{g}}{\jtermeq{\Gamma}{\ctxext{\subst{x}{P}}{\subst{x}{Q}}}{\subst{x}{\ctxext{f}{g}}}{\ctxext{\subst{x}{f}}{\subst{x}{g}}}}\\
& \inference{\jterm{\ctxext{{\Gamma}{A}}{P}}{Q}{g}\quad\jterm{\ctxext{{\Gamma}{A}}{P}}{\subst{g}{R}}{t}}{\jtermeq{\ctxext{\Gamma}{\subst{x}{P}}}{\ctxext{\subst{x}{Q}}{\subst{x}{R}}}{\subst{x}{\ctxext{g}{t}}}{\ctxext{\subst{x}{g}}{\subst{x}{t}}}}
\end{align*}
\end{lem}
Finally, when we substitute by an extended term we get an equal result as when we
substitute two consecutive times. This looks like the Curry-Howard isomorphism.

\begin{lem}
The following inference rules are valid:
\begin{align*}
& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}\quad\jfam{\ctxext{{\Gamma}{A}}{P}}{Q}}
  {\jfameq{\Gamma}{\subst{\ctxext{x}{u}}{Q}}{\subst{u}{{x}{Q}}}}\\
& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}\quad\jterm{\ctxext{{\Gamma}{A}}{P}}{Q}{g}}
  {\jtermeq{\Gamma}{\subst{u}{{x}{Q}}}{\subst{\ctxext{x}{u}}{g}}{\subst{u}{{x}{g}}}}\\
& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}\quad\jfam{\ctxext{{{\Gamma}{A}}{P}}{Q}}{R}}
  {\jfameq{\ctxext{\Gamma}{\subst{u}{{x}{Q}}}}{\subst{\ctxext{x}{u}}{R}}{\subst{u}{{x}{R}}}}\\
& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}\quad\jterm{\ctxext{{{\Gamma}{A}}{P}}{Q}}{R}{t}}
  {\jtermeq{\ctxext{\Gamma}{\subst{u}{{x}{Q}}}}{\subst{u}{{x}{R}}}{\subst{\ctxext{x}{u}}{t}}{\subst{u}{{x}{t}}}}
\end{align*}
\end{lem}
